---
title: "week-44_horror-movies"
author: "Francesco Grassi"
date: "2022-11-04"
output: 
  html_document:
    toc: true
    toc_depth: 2
    fig_retina: 1
---

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(tidyverse))
options(dplyr.summarise.inform = FALSE)
```

# Aim

Play around and make some nice plot out of the big data-set of ~35K horror movies from The Movie Database.  
Here's the plan:

- Prepare the data
- Plot top movies per decade, based on average rating
- Do some sentiment analysis, to plot the sentiments elicited by movie titles across the years

# Libraries
```{r, message=FALSE}
library(tidytuesdayR)
library(tidyverse)
library(tidytext)
library(stringr)
library(lubridate)
library(colorspace)  # modify colors
library(scales)  # custom axis scales
library(showtext)  # customize font in plots

showtext_auto()
```

# Plot settings

Let's define some settings to keep plots coherent:
```{r}

# Add some horror fonts to use in plots:
sysfonts::font_add(family = "The Truth Will Out", regular = "truthwillout.ttf")
sysfonts::font.add(family = "Blackwood Castle", regular = "Blackwood Castle.ttf")

# Size of plot text:
title_size <- 60
axis_title_size <- 30
axis_label_size <- 30
legend_title_size <- 22
legend_text_size <- 12
strip_size <- 22  # text size for facet labels

# Size of output images:
figure_width <- 20
figure_height <- 14
figure_unit <- "in"
figure_dpi <- 300

```

## Palettes

Let's also define some nice color palette.  

### Release decades

First, let's define a palette for the release decades, to use in our first plot. Here I'm starting from beautifully horror [sleepover](https://www.color-hex.com/color-palette/58166) palette by [Bluetris](https://www.color-hex.com/member/bluetris). The original palette has 5 colors. Let's interpolate to at least 8, to cover e.g. the release decades (50s-2020s) of the data-base.

```{r}
palette_decades <- colorRampPalette(c("#162e2e", "#362e2e", "#562e2e", "#762e2e", "#962e2e"), bias = 0.8)(8)
show_col(palette_decades)
```

### Sentiment 

For our second plot, we can define a different color for each sentiment to plot. We're going to use the "nrc" lexicon, which has 10 sentiments, of which we will use 8 (we won't use the general "negative" and "positive"). Even here I start from another beautiful palette by Bluetris, this time the [too bad](https://www.color-hex.com/color-palette/58926), to which I added other three colors from [Coolors](https://coolors.co/) to get to 8:
```{r}
palette_sentiment <- c("#3e437b", "#9e5577", "#a94646", "#D5BBB1", "#C6A15B", "#e3e517", "#6ecb95", "#606C38")
show_col(palette_sentiment)
```

# Prepare data

## Load data
```{r, message=FALSE}

# tuesdata <- tidytuesdayR::tt_load(2022, week = 44)  # Load datasets
# raw_data <- tuesdata$horror_movies
# write_csv(raw_data, file = "data/raw-data.csv")
raw_data <- read_csv(file = "data/raw-data.csv")

```

## Clean and prepare data

First, we will focus only on movies which have been released, so we will filter them for "status".  
We also want to extract the month, year and the decade each movie was released from the release data variable. This might turn useful in many plots

```{r}
main_df <- raw_data %>% 
  filter(status == "Released") %>% 
  mutate(release_month = month(release_date),
         release_year = year(release_date),
         release_decade = year(release_date) - year(release_date)%%10)  # find decade (maybe not the most elegant way)
```

# Top rated movie by decade

Here we want to create an horizontal bar plot, indicating the highest rated movie per decade. We can add the title of the movie inside the bar, and maybe the value of the rating right inside the bar.

We should keep in mind that many movies have a very high average score, but were rated by only one or very few people. This would lead to very distorted results. Therefore, let's select the highest rated movies only among the top 1% by vote count within each decade. Moreover, we also have to take care of ties. What we can do, (maybe not the most elegant one) is to select the top 1% movies by vote count within each decade. Then, order the movies by vote average, and *within* it, by vote count. This way, by selecting the first row per decade, we'll select the highest rated movie and will break ties based on highest vote count.

```{r}
top_plot <- main_df %>% 
  group_by(release_decade) %>% 
  filter(vote_count > quantile(vote_count, .99)) %>%  # select top 1%
  arrange(desc(vote_average), desc(vote_count), .by_group = TRUE) %>%  # "arrange()" must be told explicitly to keep groups into account 
  slice_head(n = 1)  # select top movie
```

## Ploting

Now we can plot the data:
```{r, fig.width=20, fig.height=14}
top_plot %>% mutate(title = factor(title, levels = unique(title)),  # prevent movies to be arranged by title
             release_decade = factor(release_decade, levels = unique(release_decade))) %>%  # factorize release decade to arrange it afterward
  ggplot(aes(x = fct_rev(release_decade), y = vote_average)) +  # use "fct_rev" to organize decade in ascending order
  geom_col(fill = palette_decades) +
  geom_text(aes(y = 0.2, label = title), hjust = 0, color = "white", family = "Blackwood Castle", size = axis_title_size/.pt) +  # add movie titles
  geom_text(aes(label = vote_average), hjust = 1.7, color = "white", family = "Blackwood Castle", size = axis_title_size/.pt) +  # add rating
  scale_y_discrete(expand = c(0, 0.05)) +  # reduce space between decade labels and bars
  coord_flip() +
  labs(title = "Highest Rated Horror Movie per Decade",
       subtitle = "Data from themoviedb.org",
       y = "Average Score") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = title_size, family = "Blackwood Castle"),
    plot.subtitle = element_text(size = axis_title_size, family = "Blackwood Castle"),
    axis.title.x = element_text(size = axis_title_size, family = "Blackwood Castle"),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = axis_label_size, family = "Blackwood Castle", color = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    ,plot.margin = margin(20, 20, 10, 10)  # adjust margins of plotting area
  )

showtext_auto(FALSE)
```

And finally save the plot:
```{r}
ggsave(file = "img/top-movie-by-decade.png", dpi = figure_dpi, width = figure_width, height = figure_height, units = figure_unit, bg = "white")
```

# Sentiment Analysis

Let's now do some sentiment analysis on the movie titles. We're going to use the **"nrc"** lexicon, which categorizes words into ten positive and negative emotions. Of these ten, we're going to use 8, since the other two are general "positive" and "negative", and I feel they wouldn't be too informative.

## Tokenization

First of all, we must tokenize the "title" column of our main dataset into individual words, which we'll then use for the sentiment analysis.  
We also want to remove stop-words from our analysis:
```{r}
token_df <- main_df %>% 
  select(id, title, release_year) %>% 
  unnest_tokens(word, title) %>% 
  anti_join(stop_words)
```

## Categorize words

Now that we have our tokenized data.frame, we can categorize each word by sentiment.

First of all, get the sentiment lexicon and remove "positive" and "negative" sentiments:
```{r}
nrc_lexicon <- get_sentiments("nrc") %>% 
  filter(!(sentiment %in% c("positive", "negative")))
```

Now we can use "inner_join" to retain from our tokenized df only the words that are shared with the lexicon:
```{r}
sentiment_df <- token_df %>% 
  inner_join(nrc_lexicon)
```

## Summarize by year

Let's count sentiment by year. This is in the end what we want to plot: how the presence of each sentiment in movie titles changed across time.  
We might also want to account for the fact that the total number of released movies increased across the years. If we don't, the sentiment count will be influenced by the number of released movies. One way to account for this might be to normalize the count over the number of movies released in each year:
```{r}
sentiment_by_year_df <- sentiment_df %>% 
  group_by(release_year) %>% 
  mutate(tot_year = n()) %>%  # count number of movies per year
  ungroup() %>% 
  group_by(release_year, sentiment) %>% 
  summarize(sentiment_count = n(),  # count sentiment occurrences per year
            tot_year = unique(tot_year),  # keep total number of movies per year
            sentiment_count_norm = sentiment_count/tot_year)  # normalize sentiment count over number of movies
```

## Plotting

It's finally time to plot! We want to plot the number of appearances of each sentiment in the movie titles across the years. We can go for a line plot, with years on the x-axis and count of sentiments on the y-axis.

**NOTE**: after some quick preliminary plotting, it appears that although the normalization step above, data from the earlier movies still shows a great deal of variance, especially a huge peak in the "fear" sentiment in 1953, which flattens the rest of the plot. Just to make the final figure more appealing, for the moment we **focus only on the movies from 2000 on**:
```{r}
sentiment_plot <- sentiment_by_year_df %>% 
  filter(release_year >= 2000)
```

First, we want factorize the "sentiment" variable in our df, and order the levels to match the palette chosen for this plot (this is totally arbitrary, I know):
```{r}
sentiment_plot <- sentiment_plot %>% 
  mutate(sentiment = factor(sentiment, levels = c("disgust", "sadness", "anger", "fear", "surprise", "anticipation", "trust", "joy")))
```

Let's now create the plot
```{r, fig.width=20, fig.height=14}
sentiment_plot %>% 
  ggplot(aes(x = release_year, y = sentiment_count_norm, color = sentiment)) +
  geom_line(size = 3) + 
  scale_color_manual(values = palette_sentiment) + 
  labs(title = "Sentiments Elicited by Horror Movie Titles",
       subtitle = "Data from themoviedb.org",
       x = "Year",
       y = "Percentage of Titles") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = title_size, family = "Blackwood Castle"),
    plot.subtitle = element_text(size = axis_title_size, family = "Blackwood Castle"),
    axis.title = element_text(size = axis_title_size, family = "Blackwood Castle"),
    #axis.title.y = element_blank(),
    #axis.text.x = element_blank(),
    axis.text = element_text(size = axis_label_size, family = "Blackwood Castle", color = "black"),
    panel.grid.major = element_blank(),
    #panel.grid.minor = element_blank(),
    plot.margin = margin(20, 20, 10, 10)  # adjust margins of plotting area
  )

showtext_auto(FALSE)
```