---
title: "week-44_horror-movies"
author: "Francesco Grassi"
date: "2022-11-04"
output: 
  html_document:
    toc: true
    toc_depth: 2
    fig_retina: 1
---

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(tidyverse))
options(dplyr.summarise.inform = FALSE)
```

# Aim

Play around and make some nice plot out of the big data-set of ~35K horror movies from The Movie Database.  
Here's the plan:

- Prepare the data
- Plot top 3 movies per decade, based on average rating

# Libraries
```{r, message=FALSE}
library(tidytuesdayR)
library(tidyverse)
library(tidytext)
library(stringr)
library(lubridate)
library(colorspace)  # modify colors
library(scales)  # custom axis scales
library(showtext)  # customize font in plots

showtext_auto()
```

# Plot settings

Let's define some settings to keep plots coherent:
```{r}

# Add some horror fonts to use in plots:
sysfonts::font_add(family = "The Truth Will Out", regular = "truthwillout.ttf")
sysfonts::font.add(family = "Blackwood Castle", regular = "Blackwood Castle.ttf")

# Size of plot text:
title_size <- 60
axis_title_size <- 30
axis_label_size <- 30
legend_title_size <- 22
legend_text_size <- 12
strip_size <- 22  # text size for facet labels

# Size of output images:
figure_width <- 20
figure_height <- 14
figure_unit <- "in"
figure_dpi <- 300

```

Let's also define a nice color palette. Here I'm starting from beautifully horror [sleepover](https://www.color-hex.com/color-palette/58166) palette by [Bluetris](https://www.color-hex.com/member/bluetris). The original palette has 5 colors. Let's interpolate to at least 8, to cover e.g. the release decades (50s-2020s) of the data-base.

```{r}
horror_palette <- colorRampPalette(c("#162e2e", "#362e2e", "#562e2e", "#762e2e", "#962e2e"), bias = 0.8)(8)
show_col(horror_palette)
```

# Prepare data

## Load data
```{r, message=FALSE}

# tuesdata <- tidytuesdayR::tt_load(2022, week = 44)  # Load datasets
# raw_data <- tuesdata$horror_movies
# write_csv(raw_data, file = "data/raw-data.csv")
raw_data <- read_csv(file = "data/raw-data.csv")

```

## Clean and prepare data

First, we will focus only on movies which have been released, so we will filter them for "status".  
We also want to extract the month, year and the decade each movie was released from the release data variable. This might turn useful in many plots

```{r}
main_df <- raw_data %>% 
  filter(status == "Released") %>% 
  mutate(release_month = month(release_date),
         release_year = year(release_date),
         release_decade = year(release_date) - year(release_date)%%10)  # find decade (maybe not the most elegant way)
```

# Plots

## Top rated movie by decade

Here we want to create an horizontal bar plot, indicating the highest rated movie per decade. We can add the title of the movie inside the bar, and maybe the value of therating right outside the bar.

We should keep in mind that many movies have a very high average score, but were rated by only one or very few people. This would lead to very distorted results. Therefore, let's select the highest rated movies only among the top 1% by vote count within each decade. Moreover, we also have to take care of ties. What we can do, (maybe not the most elegant one) is to select the top 1% movies by vote count within each decade. Then, order the movies by vote average, and *within* it, by vote count. This way, by selecting the first row per decade, we'll select the highest rated movie and will break ties based on highest vote count.

```{r}
p <- main_df %>% 
  group_by(release_decade) %>% 
  filter(vote_count > quantile(vote_count, .99)) %>%  # select top 1%
  arrange(desc(vote_average), desc(vote_count), .by_group = TRUE) %>%  # "arrange()" must be told explicitly to keep groups into account 
  slice_head(n = 1)  # select top movie
```

Now we can plot the data:
```{r, fig.width=20, fig.height=14}

p %>% mutate(title = factor(title, levels = unique(title)),  # prevent movies to be arranged by title
             release_decade = factor(release_decade, levels = unique(release_decade))) %>%  # factorize release decade to arrange it afterward
  ggplot(aes(x = fct_rev(release_decade), y = vote_average)) +  # use "fct_rev" to organize decade in ascending order
  geom_col(fill = horror_palette) +
  geom_text(aes(y = 0.2, label = title), hjust = 0, color = "white", family = "Blackwood Castle", size = axis_title_size/.pt) +  # add movie titles
  geom_text(aes(label = vote_average), hjust = 1.7, color = "white", family = "Blackwood Castle", size = axis_title_size/.pt) +  # add rating
  scale_y_discrete(expand = c(0, 0.05)) +  # reduce space between decade labels and bars
  coord_flip() +
  labs(title = "Highest Rated Horror Movie per Decade",
       subtitle = "Data from themoviedb.org",
       y = "Average Score") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = title_size, family = "Blackwood Castle"),
    plot.subtitle = element_text(size = axis_title_size, family = "Blackwood Castle"),
    axis.title.x = element_text(size = axis_title_size, family = "Blackwood Castle"),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = axis_label_size, family = "Blackwood Castle", color = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    ,plot.margin = margin(20, 20, 10, 10)  # adjust margins of plotting area
  )

showtext_auto(FALSE)
ggsave(file = "img/top-movie-by-decade.png", dpi = figure_dpi, width = figure_width, height = figure_height, units = figure_unit, bg = "white")
```

