---
title: "simpsons_characters"
author: "Francesco Grassi"
date: "2022-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk(echo = TRUE)
suppressPackageStartupMessages(library(tidyverse))
options(dplyr.summarise.inform = FALSE)
```


# Aim

This is just an exercise in data wrangling and plotting using the dataset from **Week 33** of the **tidytyesday** project. The dataset comes from the [Open-Source Psychometric Project](https://openpsychometrics.org/), courtesy of [Tanya Shapiro](https://twitter.com/tanya_shapiro/status/1558936093390774272?s=20&t=tI4nccuwEG4SjWd3H1Suew). Details about the dataset can be found at the [tidytuesday page](https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-08-16).

In this script I'm going to play with **personality/psychometric stats** of **The Simpsons characters**

# Libraries

```{r}

library(tidytuesdayR)
library(tidyverse)

```

# Load data

From the tidytuesday website, week 33 dataset contains three elements:

- characters: High level information about characters. Includes a notability score and links to related pages
- psych_stats: Personality/Psychometric Stats per character
- myers_briggs: Users who took the personal personality assessment tests were subsequently asked to self-identify their Myers-Briggs types. Dataset contains results

Load and extract the three elements:
```{r, eval=FALSE}

tuesdata <- tt_load(2022, week = 33)
df_char <- tuesdata$characters
df_psych <- tuesdata$psych_stats
df_myers <- tuesdata$myers_briggs
rm(tuesdata)

```

